Prestige bias in cultural evolutionary dynamics

### 2023-02-07 ms/ms.tex
revised reduction of words

\paragraph{Prestige in the literature.}

According to \citet{animal_leadership}, there are two main approaches to group decision making in nature: leadership and consensus.
Leaders would usually be high-ranking members of the group: elders, individuals with many kin relations, or individuals possessing other dominant traits.
\citet{animal_leadership} describe benefits for the closest associates of a dominant baboon, such as protection from predators. 
In some species, like the females of \emph{Eulemur fulvus rufus} (Red lemur), leaders may arise due to nutritional needs, and not due to possessing superior traits \citep{lemurs}. 
In humans, leadership also has its costs and benefits. Leaders can make decisions that would most benefit them and their closest followers, while still maintaining group cohesion. 
However, wrong decision making that harms the group could result in negative effects for the leader. 
In modern society, many humans strive for prestigious positions, as they may reap rewards greater than the risk and costs to achieve them, or due to individual personality and pressure or education from the family.

\citet{prestige_evolution} suggested that there are two types of leadership: prestige-based and dominance-based leadership.
In the latter, social status is acquired through aggression, intimidation and violence. It is also more common than prestige in non-human animals.
Their definition of prestige seems to overlap with ours. They suggest prestige is composed of estimation in the eyes of other people (similar to our success bias) and commanding position in people's minds, i.e, the number of followers other people think one has, which they define as \emph{influence}, similar to our influence bias.
\citet{prestige_evolution} show that prestige can evolve via natural selection as an efficient process to extract reproductive benefit from the flow of socially transmitted information. They back up their claims by creating testable predictions and using evidence from the literature and experiments of social sciences.
Simply put, prestige can naturally evolve when social learning already exists to save costs of individual learning.

Furthermore, according to \citet{no_replication}, the process of cultural evolution doesn't require accurate replication of cultural traits, addressed as "representation" in their paper.
They base their assumption on three claims: mental representations are non-discrete, cultural transmission is highly inaccurate, and mental representations are not replicated, but rather are `reconstructed' through an inferential process that is strongly affected by cognitive `attractors.' They describe three different models to support their points.
We see a high similarity between their model and ours. Like them, we treat the cultural trait as non-discrete (in the main model, before simplifying it to facilitate analysis). We also assume error in estimation, such that copiers do not precisely replicate their role-models, but rather reconstruct them to create a potentially different trait, which is their representation of the role-model's trait. In addition, the inferential process that they describe as strongly affected by cognitive attractors may include our definition of influence bias.

\iffalse
\paragraph{Empirical evidence of prestige bias}
\citet{prestige_cultural_learning} report the first direct tests in children that suggest the existence of prestige bias, defined as the tendency to learn from individuals to whom others have preferentially attended, learned, or deferred.
Their definition of prestige is similar to our influence bias. They showed that the odds of 3-4 years-old children learning from an adult role-model to whom bystanders had previously preferentially attended for 10 seconds were more than twice those of their learning from a role-models whom bystanders ignored.
They also note that prestige effects are domain sensitive: they found that prestigious role-models were attended more when demonstrating artifact use, whereas role-models presenting food preferences had less attendants, suggesting that the domain itself (artifact use vs. food preference) can affect the attendance, and hence the prestige of the role-model.
This lead \citet{prestige_cultural_learning} to suggest that when the trait is costly to learn individually, prestige will have a stronger bias.
It would be interesting to include costs in our model to try and observe these effects and dynamics in a large population.

@@

\paragraph{Further work.}
One path forward is an analysis of the dynamics of the adaptive success-bias weight model, in which every copier chooses its $\alpha$. It would be interesting to see the if the mean estimation error and the adaptive weight, $\alpha^*$, are converging to specific values, and how they are affected by the model parameters.
It may also be possible to relax the assumptions required for our approximations, such as homogeneous estimation error and success-bias weight.
Lastly, it would be interesting to analyze the continuous model and determine how much it differs from the dichotomous model.

@@

\fi
\paragraph{Conclusions.}
Here, we studied a model of cultural evolution under two transmission biases: the commonly studied success bias, together with influence bias, which has so far received less attention. We found approximations for this complex dynamics. We then showed that success bias affects the evolutionary dynamics much like natural selection does, whereas influence bias has a similar effect to random genetic drift. We also find a clear advantage to individuals that can choose the relative weight of the two biases.

>>>>

%\paragraph{Empirical evidence of prestige bias}
\citet{prestige_cultural_learning} report the first direct tests in children that suggest the existence of prestige bias, defined as the tendency to learn from individuals to whom others have preferentially attended, learned, or deferred.
Their definition of prestige is similar to our influence bias. They showed that the odds of 3-4 years-old children learning from an adult role-model to whom bystanders had previously preferentially attended for 10 seconds were more than twice those of their learning from a role-models whom bystanders ignored.
They also note that prestige effects are domain sensitive: they found that prestigious role-models were attended more when demonstrating artifact use, whereas role-models presenting food preferences had less attendants, suggesting that the domain itself (artifact use vs. food preference) can affect the attendance, and hence the prestige of the role-model.
This lead \citet{prestige_cultural_learning} to suggest that when the trait is costly to learn individually, prestige will have a stronger bias.
It would be interesting to include costs in our model to try and observe these effects and dynamics in a large population.

@@

%\paragraph{Further work.}
One path forward is an analysis of the dynamics of the adaptive success-bias weight model, in which every copier chooses its $\alpha$. It would be interesting to see the if the mean estimation error and the adaptive weight, $\alpha^*$, are converging to specific values, and how they are affected by the model parameters.
It may also be possible to relax the assumptions required for our approximations, such as homogeneous estimation error and success-bias weight.
Lastly, it would be interesting to analyze the continuous model and determine how much it differs from the dichotomous model.

@@

\paragraph{Conclusions.}
Here, we studied a model of cultural evolution under two transmission biases: the commonly studied success bias, together with influence bias, which has so far received less attention. We found approximations for this complex dynamics. We then showed that success bias affects the evolutionary dynamics much like natural selection does, whereas influence bias has a similar effect to random genetic drift. We also find a clear advantage to individuals that can choose the relative weight of the two biases.
